---
title: "Hands-on_Ex06"
author: "Lennis"
date: "13 May 2023"
date-modified: "`r Sys.Date()`"
edit: visual
execute: 
  warning: false
  echo: true
  eval: true
---

# Ternary Plots

## Install and launching R packages

```{r}
pacman::p_load('plotly', 'tidyverse')
```

```{r}
#| eval: FALSE
require(devtools)
install_version("ggtern", version = "3.4.1", repos = "http://cran.us.r-project.org")
```

```{r}
library(ggtern)
```

## Importing the data

```{r}
#Reading the data into R environment
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv") 
```

## Preparing the data

```{r}
#Deriving the young, economy active and old measures
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)
```

## Static Ternary Plot

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()
```

```{r}
#Building the static ternary plot
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```
```{r}
devtools::unload("ggtern")
```


## Interactive Ternary Plot

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )
```

# Visual Correlation Analysis

## Install and launching R packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

## Importing the data

```{r}
wine <- read_csv("data/wine_quality.csv")
```

## pairs() method

### Basic Correlation Matrix

```{r}
pairs(wine[,1:11])
```

```{r}
pairs(wine[,2:12])
```

### Drawing lower/upper corner

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

```{r}
pairs(wine[,2:12], lower.panel = NULL)
```

### Including correlation coefficients

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

## ggcormat() method

### Basic Correlation Matrix

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11)
```

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
```

## Building multiple plots

```{r}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

## corrplot()

```{r}
wine.cor <- cor(wine[, 1:11])
```

### Basic plot

```{r}
corrplot(wine.cor)
```

### Working with visual geometrics

```{r}
corrplot(wine.cor, 
         method = "ellipse") 
```

### Working with layout

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower")
```

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower",
         diag = FALSE,
         tl.col = "black")
```

### Working with mixed layout

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

### Combining corrgram with significant tests

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
```

```{r}
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

### Reordering corrgram

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```

### Reodering using hclust

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```

# Heatmap

## Install and launching R packages

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

## Importing and preparing the data

```{r}
wh <- read_csv("data/WHData-2018.csv")
row.names(wh) <- wh$Country
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

## Static heatmap

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

Normalize matrix column-wise

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

## Interactive heatmap

### heatmaply

```{r}
heatmaply(mtcars)
```

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```

### Data transformation

Scaling method

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

Normalizing method

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```

Percentizing method

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

### Clustering algorithm - manual

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

### Clustering algorithm - statistical

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

## Seriation

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```

## Working with colour palettes

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

## Finishing touch

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```

# Parallel Coordinates Plot

## Installing and launching R packages

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

## Preparing the data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## Static Parellel Coordinates Plot

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)
```

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

## Interactive Parallel Coordinates Plot - parallelPlot

### Basic
