---
title: "Hands-on Exercise 2"
author: "Lennis"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if ggrepel, patchwork, ggthemes, hrbrthemes and tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv", show_col_types = FALSE)
```

# Beyond ggplot2: ggrepel

Annotation is challenging, especially with large amounts of data. The chart below gives an example of an ineffective annotated chart using ggplot2. 

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

ggrepel is an extension of the ggplot2 package which provides geoms for ggplot2 to repel overlapping text as in our examples on the right. We simply replace geom_text() by geom_text_repel() and geom_label() by geom_label_repel.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

# Beyond ggplot2 themes

## ggthemes

ggplot2 comes with eight built-in themes, they are: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void(). ggthemes provides ‘ggplot2’ themes that replicate the look of plots by well-known websites such as The Economist and The Wall Street Journal. The example below shows one with The Economist theme. 

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

## hrbrthems

hrbrthemes package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

# Beyond a Single Graph: patchwork Methods

Let's first create 3 charts. 

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
```

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

```{r}
p1 + p2
```
The syntax of patchwork is simple!

```{r}
(p1 / p2) | p3
```

Autotagging can be done with patchwork. 

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

We can also create a figure with insert.

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```